---
title: "Module6"
author: "Gabi Rivera"
date: "2022-12-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Science Using Python and R: Chapter 10

For the following exercises, work with the white_wine_training and white_wine_test data sets. Use either Python or R to solve each problem.

### 11. Input and standardize both the training and test data sets.

```{r}
wine_train = read.csv('white_wine_training', sep = ',')
wine_test = read.csv('white_wine_test', sep = ',')

#Subset predictors: 
X = subset(wine_train, select = c("alcohol", "sugar"))
X_test = subset(wine_test, select = c("alcohol", "sugar"))

#Standardize predictors: 
Xs = as.data.frame(scale(X))
colnames(Xs) = c("alcohol_z", "sugar_z")

Xs_test = as.data.frame(scale(X_test))
colnames(Xs_test) = c("alcohol_z", "sugar_z")
```

### 12. Run k‐means clustering on the training data set, using two clusters.

```{r}
kmeans01 = kmeans(Xs, centers = 2)

#Save cluster membership of each record as it's own variable
cluster = as.factor(kmeans01$cluster)

#Description of each cluster
Cluster1 = Xs[ which(cluster == 1), ]
Cluster2 = Xs[ which(cluster == 2), ]
```

### 13. Give the mean of each variable within each cluster and use the means to identify a "Dry wines" and a "Sweet wines" cluster.

```{r}
#Summary
summary(Cluster1)
summary(Cluster2)
```

Cluster 1 = positive value for Sweet wine

-   Dry wine, means: -0.76

-   Sweet wine, means: 0.96

Cluster 2 = positive value for Dry wine

-   Dry wine, means: 0.49

-   Sweet wine, means: -0.62

### 14. Validate the clustering results by running k‐means clustering on the test data set, using two clusters, and identifying a "Dry wines" and a "Sweet wines" cluster.

```{r}
#Validate test dataset
kmeans01_test = kmeans(Xs_test, centers = 2)
cluster_test = as.factor(kmeans01_test$cluster)
Cluster1_test = Xs[ which(cluster_test == 1), ]
Cluster2_test = Xs[ which(cluster_test == 2), ]

summary(Cluster1_test); 
summary(Cluster2_test)
```

Cluster 1 = positive value for Sweet wine

-   Dry wine, means: -0.83

-   Sweet wine, means: 0.36

Cluster 2 = positive value for Dry wine

-   Dry wine, means: 0.37

-   Sweet wine, means: -0.19
